The binary liquid model for gravity-driven microchannel simulations

### 2010-11-22 Text/paper.tex
Added half modifications.

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[numbers]{natbib}
\usepackage{color}

@@

\begin{abstract}
The motion towards exploration of nano- and microscale phenomena brings new
challenges to simulation tools. The continuous physics approach is limited on
those scales, and methods such as Molecular Dynamics (MD) or the Lattice
Boltzmann Method (LBM) become more appropriate. This work aims to
reproduce the gravity driven microchannel flow of a free surface liquid with the
help of the free-energy binary liquid lattice Boltzmann method. In the low capillary number
$\Ca$ regime, the flow is thoroughly described by Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. The binary
liquid model we use is a continuous-interface multiphase model, which presents
the question of how to properly resolve the interface in comparison with the film thickness to
obtain Bretherton correlations.
In this work we investigate how the binary liquid model parameters and the grid resolution
need to be adjusted to properly simulate microchannel flows.  The presented results
should be of interest for industrial applications and to researchers simulating
Bretherton phenomena.
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
microchannels. It was found that the front meniscus film width is proportional
to $\Ca^{2/3}$ for Capillary numbers less than $0.003$. While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-bretherton}, its simulation using diffusive interface methods
is difficult.  If the interfaec is spread over several grid nodes, the question of
proper film width resolution arises. While the discontinuous interface models
are good at simulating pheonomena in simple geometries, the continuous interface
methods are more flexible and capable of resolving problems involving coalescence
or droplet breakup. This work
concentrates on resolving the parameter range of the binary liquid lattice
Boltzmann method to validate it the Bretherton regime.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and the incorporation of
physical phenomena on the microscopic level, as is the case in multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in a continuous way. Such a representation
brings issues with the film thickness resolution versus the interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusive interface should be resolved in such a way as to have
negligible effect on the physics of the film.

This work examines the binary liquid free-energy LB model \cite{swift}. The
model simulates two liquids with the assumption of uniform overall
density. Therefore, the question of how to simulate the Bretherton problem
where the two phases have two completely different densities (i.e. the two fluids
are liquid and gas) is of interest. However, in the case of the motion of a
liquid slug and a bubble in a capillary the governing factor is the capillary number
and the ratio of the viscosities, not of the densities, is the critical parameter.

The goal of this paper is to give an answer as to how to choose the
binary liquid model parameters, the viscosities ratio and the grid
resolution, in order to obtain correct Bretherton/Taylor film thickness.
The paper is organized as follows.  First, we briefly
explain the binary liquid lattice Boltzmann model and summarize the literature
for the Bretherton problem. Then, the parameters involved in the
simulation are examined and presented in the results section. The paper is
concluded with a summary of the main findings.

\section{Bretherton bubble flow}
Bretherton in his significant work \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film width
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
it was realized \cite{wong-films,wong-pressure} that the film deposition
is proportional to $\Ca^{2/3}$ only near the front meniscus and that the film width
varies throughout the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
To be able to predict consistently the flow pattern for capillaries in
different ranges of parameters, simulations are developed and their codes
are typically validated with the small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary.
\citet{kreutzer-pressure-drop} used the finite volume method to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} in
\cite{wong-films,wong-pressure} studied three-dimensional bubbles in
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied air finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with previously published results.
The solution is available for the circular and planar cases.

@@

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discretized velocity set.
The LBE is typically implementated as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $F_i$ is the external force population. The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and velocity fields, and the
other to represent the phase field.

The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is a parameter of the free-energy model  The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
Parameters speficic to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken to depend linearly on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k A}{9}}$.

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film width.
 \item Spurious currents arising from surface tension discretization
can influence mass transfer in the thin film, thereby comoromising the
overall solution. The bubble should therefore be long enough to decrease the
influence of spurious currents.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is a single density, two
viscosities model. The classical Bretherton problem is a free-interface problem
where the film thickness is established by itself. Thus, the viscosity and density ratios for the
Bretherton problem should
be large. However, in terms of the lattice Boltzmann free-energy framework those issues are not
addressed and will be studied in the current work.
  \item The Bretherton problem addresses bubble flow driven by a pressure
difference, not a body force.  It is however relatively difficult to impose a
pressure difference in the framework of the binary liquid model, as two
simultaneous boundary conditions with the associated gradients and laplacians are required.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is concerned.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability.
 \item Most simulations are done for a circular channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate Laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue.
Our benchmark is therefore done is for a plane-symmetric case in the two-dimensional case. We do
not address the axis-symmetric two-dimensional case since it implies the derivation of additional
mass and force terms for the square based LBM grid \cite{halliday-circular}.
\end{itemize}
Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow driven by a body force.  An analytical
solution to this problem is to the best of our knowledge not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and the
bubble shape
is different in the front and rear menisci. \citet{wong-films} indicated that the
film width for polygonal capillaries
changes from $\Ca^{1/2}$ on distances $x\propto \Ca^{-1}$ to $\Ca^{2/3}$ for
distances $x \gg \Ca^{-1}$. Therefore, for a bubble to have a formed shape in
the  front meniscus we chose the bubble length to be $5$ channel heights. The
channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions.
Figure \ref{fig:benchmark:sketch} represents a sketch of the geometry
used in the benchmark.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
We now intrduce the procedure of nondimensionalization, as well as
the initialization techniques we used to be able to predict parameters necessary for
simulation. The capillary number governs the interface thickness and is defined as:
\begin{equation}
Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
\end{equation}
where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the surface tension. The
bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
thickness is examined.
While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
or defined through the model parameters, the velocity $U_{\mathrm{bubble}}$ is not.
In the simulation we approximately match the velocity with the
Poiseuille profile center velocity, which is given by the equation:
\begin{equation}
\begin{aligned}
U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{\mathrm{bubble}}\approx
\frac{{N_y}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
By taking the Poiseuille profile one can estimate the body force through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{N_y}^2}\gamma Ca.
\end{aligned}
\end{equation}
In practice, we iterate adjusting the body force value until both the desired velocity and the
capillary numbe are achieved.

After one calculates the parameters necessary to run a simulation, one needs to
initialize the macroscopic fields and particle populations.
The velocity is initialized with $0$ everywhere. Populations are initialized
with the equilibrium values for the binary liquid model, including all the phase
gradients and Laplacians. The liquid phase is initialized with $\phi = 1$ and the
gas phase is initialized with $\phi = -1$. Though the initialization film thickness
is not involved in the final result for the front meniscus (see Fig.
\ref{fig:different:initialization:widths}) we keep the initialized film
thickness as close as possible to the already obtained numerical film
thickness values \cite{giavedoni-numerical}. This is done to minimize
the simulation time, as starting far away from the steady-state requies
additional time for the relaxation.
\begin{figure}
\includegraphics*[bb=-20 360 620 430,width=0.97\textwidth]
{Figures/Init/initbegin12.eps}
\includegraphics*[bb=-20 360 620 430,
width=0.97\textwidth]{Figures/Init/initbegin16.eps}
\includegraphics*[bb=-20 360 620 430,
width=0.97\textwidth]{Figures/Init/initbegin20.eps}
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
\includegraphics*[bb=-20 360 620
430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
\caption{The phase plots for different bubble widths as
$H_{\mathrm{eff}}-12$,$H_{\mathrm{eff}}-16$ and
$H_{\mathrm{eff}}-20$. The grid for the simulation is $102 \times 1501$. The phase
profiles were obtained after $2\cdot10^5$ time steps. One can see that even if the
system is initialized with different bubble volumes, the bubbles always relaxate
to a shape where the film thickness is conserved. For the given parameters the
interface thickness is $0.0615$, $0.0611$, $0.0605$.
\label{fig:different:initialization:widths}}
\end{figure}

>>>>

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{color}

@@

\begin{abstract}
This work shows that the free-energy binary liquid lattice Boltzmann method is able to
reproduce the gravity driven microchannel flow of an air bubble through liquid. In the low capillary
number
$\Ca$ regime (defined in Eq. \ref{capillary:number:definition}), the flow is thoroughly described by
Bretherton \cite{bretherton},
with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
from the Bretherton correlations for $Ca\geq 0.005$, studied experimentally and numerically. This
work aims to examine the lattice Boltzmann binary liquid model feasibility to obtain valid results
for an air bubble flow in liquid for a range of $Ca$ number. The binary
liquid model is a continuous-interface multiphase model, which raises
the question of how to properly resolve the interface in comparison with the film thickness.
It was found that the film interface needs to be as $50$
percent in comparison with the film thickness for simulation to give convergent results. While
the convergency criteria is fulfilled the results are of good comparison with the
numerical simulations of \citet{giavedoni-numerical} in the range of $Ca$ from $0.01$ to $1$. 
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow describes the flow of long bubbles in
microchannels. It was found that the front meniscus film width is proportional
to $\Ca^{2/3}$ for capillary numbers less than $0.003$, where the capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is a propagation bubble velocity, and
$\gamma$ is the surface tension between air and liquid.
While there are number
of works which simulate the Bretherton problem with the boundary value method
\cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
involving complex geometries, free interface motion or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for those kind of simulations.
However, if the interface is spread over several grid nodes, the question of
proper film width resolution in comparison with the interface arises.  This work
concentrates on resolving the parameter range of the binary liquid lattice
Boltzmann method to validate it in the range of capillary numbers.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries andallows for incorporation of
physical phenomena on the microscopic level, as in the case in multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface in a continuous way. Such a representation
brings issues with film thickness resolution versus interface
resolution. In other words, when it comes to the resolution of the film
thickness the diffusive interface should be resolved in such a way as to have
negligible effect on the physics of the film.

This work examines the binary liquid free-energy LB model \cite{swift}. The
model simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for air and liquid, which are of largely
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effect on the film thickness for the range of Reynolds numbers.
Therefore, the governing factor for microchannel flows is not the densities ratio, but the
viscosities ratio. Moreover, the results based on the uniform density, as in the case of the LBM
binary liquid model, are of good comparison with other simulations
\cite{giavedoni-numerical,heil-bretherton} taking into the account inertia effects.

The goal of this paper is to give an answer as to how to choose the
binary liquid model parameters, the viscosities ratio and the grid
resolution, in order to obtain correct film thicknesses.

The paper is organized as follows.  First, we briefly
explain the binary liquid lattice Boltzmann model and summarize the literature
for the Bretherton problem. Then, the parameters involved in the
simulation are examined and presented in the results section. The paper is
concluded with a summary of the main findings.

\section{Bretherton bubble flow}
Bretherton \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film width
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later on
it was realized \cite{wong-films,wong-pressure} that the film deposition
is proportional to $\Ca^{2/3}$ only in the specific region behind the front meniscus and
that the film width
varies over the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed as well a Reynolds number dependence, and
a deviation from the $\Ca^{2/3}$ rule for capillary numbers larger than $0.003$.
To be able to predict consistently the flow pattern for capillaries in
different ranges of parameters, simulations have been developed and are typically validated with the
small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary using the finite volume method.
\citet{kreutzer-pressure-drop} used the finite volume method as well to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} \cite{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and showed the derivations of different shapes in the
slug crossections and menisci appearance.
\citet{heil-bretherton,ingham-plates} studied air finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers using finite element method. \citet{giavedoni-numerical} performed crossvalidation of the
finite element solution developed by them with previously published results.
The solution is available for circular and planar cases.

@@

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implementated as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $F_i$ is the external force population. The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and velocity fields, and the
other to represent the phase field.

The model we use in this work is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
\phi \partial_y \phi \bigr), i=1\div8\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is a parameter of the free-energy model  The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
Parameters speficic to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The system can be shown to restore the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity and
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter. The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken as linearly dependant on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}\tau_{\mathrm{liq}}$.
The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k A}{9}}$.

\section{Lattice Boltzmann benchmark}
To properly simulate and to compare simulation results with the data
published in the literature, one needs to design a numerical benchmark addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive nature of the interface in the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the interface in comparison with the film width. The same applies to spurious
currents arising from surface tension discretization which
can influence mass transfer in the thin film, thereby compromising the
overall solution. Therefore the diffusive interface should be of nengligible influence on the film
thickness.
 \item The wettability coefficient which in principle can control the
dynamic angle and the menisci appearance \cite{pagonabarraga-finger} should not
influence the overall solution.
 \item The free-energy model examined in the work is a single density, two
viscosities model. The classical Bretherton problem is a free-interface problem between liquid and
air
where the film thickness is established by itself. Thus, the viscosity and density ratios for the
classical
Bretherton problem should
be large. However, in terms of the lattice Boltzmann binary liquid framework only the viscosities
ratio can be addressed. The results show that the results are consistent with other works and
inertia effects can be neglected.
  \item The Bretherton problem addresses a bubble flow driven by a pressure
difference, not a body force. However, to the best authors knowledge the pressure boundaries
conditions are still not yet developed for the binary liquid model, because it involves the
treatment of associated gradients and laplacians on the boundary.
This kind of pressure boundary conditions for the binary liquid model is currently
under development, and preliminary results show that the differences between
pressure difference flows and flows driven by a body force are small as far as the film
deposition thickness is of interest.  In this work we limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability. The case of
body force driven flows implies the periodic boundary conditions. As soon as the periodic boundary
conditions are applied, not one bubble but whole bubble train is simulated. In this case one needs
to insure that bubbles do not influence each other making large distance between them.
 \item Most simulations are done for a channel with circular crosssection,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the Laplacian and gradients near boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed specifically for curved boundaries,
the analysis of error introduced by the boundary conditions becomes an issue. Though the circular
axis-symmetric case can be simulated on rectangular grid by introducing specific mass and force
terms in the continuity and Navier-Stokes equation \cite{halliday-circular}, the lattice Boltzmann
incorporation of force and mass terms in the process of development and validation. Therefore
the benchmark is chosen as a plane-symmetric two-dimensional case.

\end{itemize}
Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow driven by a body force.  An analytical
solution to this problem is to the best of our knowledge not yet known for the range of capillary
numbers.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and the
bubble shape
is different in the front and rear menisci depending on Bond number $Bo$. \citet{wong-films}
analytically described the film thickness variation along a bubble varying from $\Ca^{1/2}$ to
$\Ca^{2/3}$. The bubble should be long enough to have the film thickness proportional to
$Ca^{2/3}$. Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
to be sufficient to give consistent with theory results. The
channel length is taken  as $3$ bubble lengths (or $15$ channel heights)
to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions.
Figure \ref{fig:benchmark:sketch} is a sketch of the geometry
used in the benchmark.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
We now introduce the procedure of nondimensionalization, as well as
the initialization techniques we used to be able to determine parameters necessary for
simulation. The capillary number governs the interface thickness and is defined as:
\begin{equation}
Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
\end{equation}
where $\mu_{\mathrm{liq}}$ is the liquid viscosity, and $\gamma$ is the surface tension. The
bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
thickness is examined.
While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
or defined through the model parameters, the bubble velocity $U_{\mathrm{bubble}}$ is not given
explicitly.
In the simulation we approximately match and initialize the velocity with the
Poiseuille profile center velocity, which is given by the equation:
\begin{equation}
\begin{aligned}
U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H_{\mathrm{eff}}^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{\mathrm{bubble}}\approx
\frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
where $N_y$ is the grid number in the vertical direction. As it will be discussed further, the
effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes.
By taking the Poiseuille profile one can estimate the body force through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma Ca.
\end{aligned}
\end{equation}
The body force (\ref{poiseuille:velocity:center}) applied to the problem will cause the bubble
velocity different from the implied by the capillary number because of the multiphase flow. That's
why to achieve the 
desired velocity and the
capillary number the iteration procedure to find the appropriate body force needs to be done.
However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
starting approximation point.

After one calculates the parameters necessary to run a simulation, one needs to
initialize the macroscopic fields and particle populations.
The velocity is initialized with $0$ everywhere. Populations are initialized
with the equilibrium values for the binary liquid model, including all the phase
gradients and Laplacians. The liquid phase is initialized with $\phi = 1$ and the
gas phase is initialized with $\phi = -1$. Though the initial film thickness
is not involved in the final result for the front meniscus (see Fig.
\ref{fig:different:initialization:widths}) we keep the initialized film
thickness as close as possible to the already obtained numerical film
thickness values \cite{giavedoni-numerical}. This is done to minimize
the convergence simulation time, as starting far away from the steady-state requies
additional time for the relaxation.
\begin{figure}
\includegraphics*[bb=-15 360 630 430,width=0.97\textwidth]
{Figures/Init/initbegin12.eps}
\includegraphics*[bb=-15 360 630
430,width=0.97\textwidth]{Figures/Init/initfinish12.eps}\\
\includegraphics*[bb=-15 360 630 430,
width=0.97\textwidth]{Figures/Init/initbegin16.eps}
\includegraphics*[bb=-15 360 630
430,width=0.97\textwidth]{Figures/Init/initfinish16.eps}\\
\includegraphics*[bb=-15 360 630 430,
width=0.97\textwidth]{Figures/Init/initbegin20.eps}
\includegraphics*[bb=-15 360 630
430,width=0.97\textwidth]{Figures/Init/initfinish20.eps}\\
\caption{The phase plots for different bubble widths as
$H_{\mathrm{eff}}-12$, $H_{\mathrm{eff}}-16$ and
$H_{\mathrm{eff}}-20$. The grid for the simulation is $102 \times 1501$. The results are rescaled
on $N_y$.The phase
profiles were obtained after $2\cdot10^5$ time steps. One can see that even if the
system is initialized with different bubble volumes, the bubbles always relax
to a shape where the film thickness is conserved. For the given parameters the
interface thickness rescaled at the effective channel width $H_{\mathrm{eff}}$  is
$0.0701$,$0.0697$,$0.0692$ measured at the dimensionless coordinate $x=13$.
\label{fig:different:initialization:widths}}
\end{figure}
